Lens Modeling in \sw

### 2014-06-16 text/partA.tex
editing

The encouraging early results prompted the question: could modeling of
the lenses also be done by volunteers?  There are several software
packages available for lens modelling \citep[for a review,
  see][]{2013NewAR..57....1L} and work has been done on generic
interfaces \citep{2014A&C.....5...28L}, lens modelling has remained
the domain of specialists.  An early proposal for what became \sw had
included a prototype lens modelling tool \citep{2010AAS...21543527N},
but the learning curve for modelling was subsequently considered too
long to attract potential volunteers.  On the other hand, some {\em
  Spacewarps\/} volunteers were quite experienced from earlier
projects, having spent a thousand hours or more with data, and very
interested in more demanding projects.  The interests of
citizen-science communities are just beginning to be studied
\citep[e.g.,][]{2013AEdRv..12a0106J}, but it is clear that some
volunteers welcome open-ended challenges, and sometimes these have led
to new scientific results: one example is the discovery of an
exceptional extra-solar planet \citep{2013ApJ...768..127S}; another is
the development of new algorithms for protein folding
\citep{Khatib22112011}.

A characteristic of strong gravitational lensing is that lensed images
do not look much like the source, and still less do they resemble the
mass distribution.  Modellers develop an intuition for image
configurations, and it is necessary to do so before one can begin to
model lenses.  Indeed, the modelling section of
\cite{1981ApJ...244..723Y} can be read as researchers developing
lens-modellers' intuition for the first time.

@@

\section{A lens modeling program}

\spl is an extension of the GLASS framework for modelling lenses
\citep{2014arXiv1401.7990C}, with two innovations.
\begin{itemize}
\item \spl is friendly to the forum style of citizen-science projects,
  without sacrificing any features of GLASS.  The program is invoked
  online, and the modelling history is logged in a transparent way,
  enabling forum discussions and incremental improvement by different
  people.  Moreover, any result can be revisited and post-processed if
  desired.  These features are of general interest for citizen-science
  projects, but somewhat outside the scope of the present paper.
\item The innovation specific to lens modelling is the use of
  ``spaghetti diagrams'' or saddle-point contours.  These were
  originally introduced to gravitational lensing by
  \cite{1986ApJ...310..568B} as a way of classifying lensed images,
  and are sometimes shown as part of the output of lens models
  \citep[for
    example][]{2001ApJ...557..594R,2003ApJ...590...39K,Lubini2012}.
  In \spl, however, spaghetti diagrams are the {\em input\/} through
  which the modeller tells the program what to do.
\end{itemize}

@@

\subsection{Fermat's Principle} \label{sec:Fermat}

@@

\subsection{``Spaghetti'' diagrams} \label{sec:arriv}

@@

\subsection{\spl} \label{sec:SpaghettiLens}

>>>>

The encouraging early results prompted the question: could modeling of
the lenses also be done by volunteers?  An early proposal for what
became \sw included a prototype lens modelling tool
\citep{2010AAS...21543527N}.  There are several software tools for
lensing modelling available, and work has been done on generic
interfaces \citep{2013NewAR..57....1L,2014A&C.....5...28L}.  Moreover,
some {\em Spacewarps\/} volunteers are quite experienced from earlier
projects, having spent a thousand hours or more with data, and very
interested in more demanding projects.  The interests of
citizen-science communities are just beginning to be studied
\citep[e.g.,][]{2013AEdRv..12a0106J}, but it is clear that some
volunteers welcome open-ended challenges, and sometimes these have led
to new scientific results: one example is the discovery of an
exceptional extra-solar planet \citep{2013ApJ...768..127S}; another is
the development of new algorithms for protein folding
\citep{Khatib22112011}.  All these are grounds for optimism.  There
is, however, a basic difficulty in strong gravitational lensing.
Lensed images do not look much like their source, and still less do
they resemble the lensing-mass distribution.  To model a lens, one
needs to either do a lot of random guessing, or have a good intuition
for what works.

In this paper, we propose a way around the difficulty, and report on a
modelling test on \sw using simulated lenses.  There are three aspects
to this work, covered in the following three sections.

In section \secref{Fermat} we introduces a theoretical innovation,
which we call a ``spaghetti diagram''.  This is a cartoon-like markup
of a lensed system, which implicitly encodes the basis of a mass
model.  A spaghetti diagram is basically the saddle-point contours
originally introduced to gravitational lensing by
\cite{1986ApJ...310..568B} as a way of classifying lensed images.
They are sometimes shown as part of the output of lens models
\citep[for
  example][]{2001ApJ...557..594R,2003ApJ...590...39K,Lubini2012}.  In
the present work, however, spaghetti diagrams are the {\em input\/}
through which the modeller tells the program what to do.

In section \secref{SpaghettiLens} we describe the \spl program, which
implements everything.  \spl is an extension of the GLASS framework
for modelling lenses \citep{2014arXiv1401.7990C}.  \spl is designed to
be friendly to the forum style of citizen-science projects, without
sacrificing any features of GLASS.  The program is invoked online, and
the modelling history is logged in a transparent way, enabling forum
discussions and incremental improvement by different people.
Moreover, any result can be revisited and post-processed if desired.
These features are of general interest for citizen-science projects,
but somewhat outside the scope of the present paper.

@@

\section{Fermat's principle and spaghetti diagrams} \label{sec:Fermat}

@@

\subsection{Geometrical and gravitational time delays}

@@

\subsection{Arrival-time contours} \label{sec:arriv}

@@

\section{A lens-modelling program} \label{sec:SpaghettiLens}
